Even-point Multi-loop Unitarity and its Applications: Exponentiation, Anomalies and Evanescence

### 2023-07-03 eventPointTwoLoop.tex

\documentclass[12pt,letter]{article}
\usepackage{jheppub}
\usepackage{amsmath,amssymb}
\usepackage[dvipsnames]{xcolor}
%\usepackage{hyperref}
\usepackage{xspace}
\usepackage{ifdraft}
\usepackage{epstopdf}
\usepackage{slashed}
\usepackage{diagbox}
\usepackage{colortbl}
\usepackage{shuffle}
\usepackage{bbold}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{stackengine}
\def\RA{\rlap{\scalebox{1.6}{$\leftrightarrow$}}}
\def\DA{\smash{\bclap{\scalebox{1.6}{$\downarrow$}}}}
\def\mystrut{\rule[-2ex]{0ex}{6ex}}

\usepackage{tikz-feynman} 
\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={hred},
    citecolor={hgreen},
    urlcolor={hblue}
}
\tikzset{graviton/.style={decorate, decoration={snake, amplitude=.4mm, segment length=1.5mm, pre length=.5mm, post length=.5mm}, double}}
\tikzfeynmanset{compat=1.1.0}
\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\usetikzlibrary{external}  
          %% Load the `external` library
\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\tikzexternalize[                     %% Activate externalization
  prefix=images/,                    %% Avoid cluttering the directory
 system call={                       %% Use lualatex in system call
    pdflatex \tikzexternalcheckshellescape -halt-on-error -shell-escape -interaction=batchmode -jobname="\image" "\texsource" || rm "\image.pdf"
  },
]

@@

\definecolor{cutred}{RGB}{219,56,49}
\definecolor{hgreen}{RGB}{25,176,146}
\definecolor{hgreen1}{RGB}{175,230,175}
\definecolor{hblue}{RGB}{52,152,219}
\definecolor{hbluedark}{RGB}{36, 106, 160}
\definecolor{hblue1}{RGB}{255,255,166}
\definecolor{hred}{RGB}{216,83,117}
\definecolor{hreddark}{RGB}{151, 58, 81}
\definecolor{hred1}{RGB}{255,155,155}
\definecolor{cutred}{RGB}{219,56,49}
\definecolor{hgrey4}{RGB}{75,75,75}
\definecolor{hgrey5}{RGB}{50,50,50}
\definecolor{hgrey3}{RGB}{100,100,100}
\definecolor{hgrey}{RGB}{125,125,125}
\definecolor{hgrey2}{RGB}{125,125,125}
\definecolor{hgrey1}{RGB}{150,150,150}
\definecolor{hgrey0}{RGB}{175,175,175}
\definecolor{darkgreen}{RGB}{59,126,108}

@@

\def\tree{{\rm tree}}
\def\pol{\varepsilon}
\def\Tr{\, {\rm Tr}}
\def\tr{\, {\rm tr}}
\def\eps{\varepsilon}
\def\e{\varepsilon}
\def\ep{\varepsilon}
\def\SYM{MSYM}
\def\nn{\nonumber}
\def\sec#1{section~\ref{#1}}
\def\eqn#1{eq.~(\ref{#1})}
\def\Eqn#1{Equation~(\ref{#1})}
\def\eqns#1#2{eqs.~(\ref{#1}) and~(\ref{#2})}
\def\Eqns#1#2{Eqs.~(\ref{#1}) and~(\ref{#2})}
\def\Figref#1{Fig.~\ref{#1}}
\def\secref#1{section~\ref{#1}}
\def\Neqfour{{{\cal N}=4}}
\def\NeqFour{{{\cal N}=4}}
\def\Neqeight{{{\cal N}=8}}
\def\NeqEight{{{\cal N}=8}}
\def\NeqOne{{{\cal N}=1}}
\def\Fact{{\cal F}}
\def\f{\widetilde f}
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\bea{\begin{eqnarray}}
\def\eea{\end{eqnarray}}
\def\ba{\begin{eqnarray}}
\def\ea{\end{eqnarray}}
\def\Ksl{{\s K}}
\def\ksl{\s{k}}
\def\Perm{{\cal P}}
\def\M{{\cal M}}
\def\ve{\varepsilon}
\def\tlambda{{\tilde\lambda}}
\def\MHVbar{$\overline{\hbox{MHV}}$}
\def\NMHVbar{$\overline{\hbox{NMHV}}$}
\def\P{{\rm P}}
\def\NHP{{\rm NHP}}
\def\mud{\lambda}
\def\bowtie{{\rm bow\mbox{-}tie}}
\newcommand{\cred}{\bf \color{red}}
 \newcommand{\cblue}{\color{blue}}
 \definecolor{MattOrange}{rgb}{1.0,0.4,0.2}
\newcommand{\cob}{ \bf \color{MattOrange}}
\newcommand{\andd}{\ , \quad \text{and}  \quad}
\newcommand{\forr}{\ , \quad \text{for}  \quad}

@@

\title{Insights of Even-point Multi-loop Unitarity: Exponentiation, Anomalies and Evanescence}

@@

\section{Introduction}\label{sec:intro}
Recent decades have seen  significant advances in our ability to compute scattering amplitudes to higher orders in perturbation theory. At the epicenter of this explosion in the generation of sharp $S$-matrix data is the unitarity method \cite{UnitarityMethod, Fusing, BDKUniarityReview}, which bypasses the standard Feynman diagram approach by constructing higher order amplitudes directly from lower order on-shell information. At one-loop, the unitarity method allows one to extract amplitudes by directly computing a series of unitarity cuts \cite{Forde:2007mi}. On-shell methods have furthermore unveiled novel amplitude-level structure, like the  duality between color and kinematics \cite{BCJ} and associated double-copy construction \cite{BCJLoop}, which are frequently obscured by the traditional classical Lagrangian description of the theories. While much of the research in perturbative calculations thus far has focused on gauge theory and gravity, recent literature has investigated to what extent the $S$-matrix of effective field theory is constrained by on-shell data \cite{Adams:2006sv,Cheung:2015ota,Cheung:2016drk,Carrasco:2019yyn,Arkani-Hamed:2020blm,Carrasco:2021ptp,Chi:2021mio,Bonnefoy:2021qgu,Carrasco:2022lbm,Carrasco:2022sck,Green:2022slj,Pavao:2022kog,Chen:2022shl,Chen:2023dcx,Brown:2023srz}.

@@

\section{Review} 
\label{sec:Review}
\subsection{Color-dressed and ordered amplitudes}
Here we provide an overview of the amplitudes nomenclature and organizational principles we use throughout the work. When working with scattering amplitudes, $\mathcal{A}$, in an on-shell framework it is convenient to introduce the following graphical description at general multiplicity, $n$, and loop order, $L$,
\be\label{fullAmp}
\mathcal{A}_{n,L} = \int \prod_{i=1}^L\frac{d^D l_i}{(2\pi)^D} \sum_{g} \frac{1}{S_g}\frac{\mathcal{N}_g}{D_g}
\ee
As written above, $S_g$ are the internal symmetry factors for a Feynman diagram, $g$, the propagator structure is captured by the denominator function, $D_g$, and the theory-dependent interaction vertices determine numerator functions, $\mathcal{N}_g$. As written the numerator functions carry all non-propagator kinematic information and any color-data that would be encoded in the interaction Feynman rules of the theory. By color-data we mean the typical weighting of feynman diagrams by the representation of the particles described.  Most of the theories we describe are defined in the adjoint, so the color-data involves the usual dressing of vertices with  antisymmetric structure constants, $f^{abc}$.  Although, as we will see, it will be useful to describe certain counterterms in terms of symmetric $d^{abc}$ color-weights,

Fo  gauge theories, it is often useful to further decompose the full, or color-dressed, amplitudes of \eqn{fullAmp} into purely kinematic building blocks where the color information is stripped away. For example, at tree level, gauge theory $n$-point amplitudes can be re-expressed in terms of a trace basis decomposition as follows:
\be
\mathcal{A}_{n,\text{tree}} = \sum_{\sigma \in S^{n-1}} \text{Tr}(T^1 T^{\sigma(2)}\cdots T^{\sigma(n)}) A_n(1,\sigma(2),...,\sigma(n))
\ee
where the sum is taken over the inequivalent orderings of group theory generator traces. A similar decomposition applies to one-loop gauge theory amplitudes, which we will describe in \sect{sec:Loops}. The color-ordered functions, $A_n$, are called {partial amplitudes}, as they only contain on-shell information for a particular color ordering. However, like full amplitudes, they factor on poles to lower-point partial amplitudes. For vector theories partial amplitudes are independently gauge invariant since they weight linearly independent color traces.

When expressed in this form, many gauge theories reveal hidden redundancy that is obscured by the color-dressed amplitude formulation of \eqn{fullAmp}. The $(n-1)!$ partial amplitudes for a large catalog \cite{BCJreview} of adjoint gauge theories, including Yang-Mills and NLSM, are linearly related to a smaller set of basis amplitudes. The first such set of partial amplitude relations identified by Kleiss and Kluijf \cite{KK}, known formally as KK relations, relates amplitudes different color orderings and signature to a basis of $(n-2)!$ partial amplitudes:
\be\label{KK}
A(1,\alpha,n,\beta)=(-1)^{|\alpha|}\sum_{\sigma\in \alpha \shuffle \beta^T} A(1,\sigma,n)
\ee
where $\beta^T$ denotes the inverse ordering of $\beta$, and $\shuffle$ is the shuffle product that runs over ordered permutations of $\alpha$ and $\beta^T$. More recently, Bern, Johansson and one of the authors, (BCJ) further identified a set of kinematic relations that reduced the size of $n$-point partial amplitude basis down to $(n-3)!$ via BCJ relations \cite{BCJ}:
\be\label{BCJ}
\sum_{i=2}^{n-1}k_1\cdot(k_2+\cdots +k_i) A(2,...,i,1,...,n)=0
\ee
Underlying both sets of amplitude relations are hidden graphical  principles that manifest the amplitude level redundancy of \eqn{KK} and \eqn{BCJ}, which we  describe in the next section. Before moving on, we emphasize that everything described thus far applies in arbitrary spacetime dimensions. As such, the aforementioned nomenclature and organization are well suited for constructing loop integrands compatible with  dimensional regularization.
\label{subsec:AmpReview}

\subsection{Color-Kinematics Duality and the Double-Copy}
\label{subsec:DCReview}
To apprehend this graph-based structure for color-dressed theories, we first rewrite the full amplitude of \eqn{fullAmp} as a sum over {cubic graphs}, as follows:
\be\label{fullAmpCubic}
\mathcal{A}_{n,L} = \int \prod_{i=1}^L\frac{d^D l_i}{(2\pi)^D} \sum_{g\in \Gamma^{(3)}_{n,L}} \frac{1}{S_g}\frac{C_gN_g}{D_g}
\ee
where $N_g$ and $C_g$ are the kinematic numerators and color factors, respectively, and $ \Gamma^{(3)}_{n,L}$ denotes the set of cubic diagrams at $n$-point $L$-loop. In this formulation of the amplitude, contact diagrams are absorbed into cubic graphs by multiplying by inverse propagators. The procedure of assigning contact diagrams to cubic graphs is referred to as {generalized gauge freedom} \cite{BCJ}. In practice, there are many different ways to assign contact diagrams to cubic graphs.

The gauge theories that we study in this paper have sufficient generalized gauge freedom to assign contact diagrams to kinematic numerators such that at every multiplicity at tree level one can write the amplitude in terms of $N_g$ that obey the same algebraic relations as the color factors, $C_g$. For color factors that are composed of adjoint structure constants, $f^{abc}$ they must be anti-symmetric around vertices, and be related by Jacobi relations about each edge.   For $SU(N_c)$ we can take $f^{abc} \propto \text{Tr}(T^a[T^b,T^c])$ for arbitrary representation $T$ of the gauge group.
\iffalse
Explicitly, given the internal edge of a cubic graph, $\mathcal{G}{(ij|kl)}$, described functionally in terms of external edges,
\be
\mathcal{G}{(ij|kl)}=\threeTree{$j$}{$k$}{$l$}
\ee
we can construct color factors, $C(ij|kl)$, and kinematic factors, $N(ij|kl)$, that obey the following set of functional relations on all vertices of internal edges:
\begin{itemize}
\item \textbf{Antisymmetry Relations} $\Leftrightarrow$ \textbf{KK Relations}
\be
C{(ij|kl)}+C{(ij|lk)} =0  \qquad \qquad N{(ij|kl)}+N{(ij|lk)} =0 
\ee
\item \textbf{Jacobi Relations} $\Leftrightarrow$ \textbf{BCJ Relations}
\be
C{(ij|kl)}+C{(il|jk)} +C{(ik|lj)} =0  \qquad N{(ij|kl)}+N{(il|jk)} +N{(ik|lj)} =0 
\ee
\end{itemize}
\fi
We will also consider generalization beyond the adjoint to color-weights that can include $d^{abc}$ and associated algebraic relations.   For $SU(N_c)$ we can take $d^{abc} \propto Tr(T^a\{T^b,T^c\})$.  Symmetric-structure color obey symmetric algebraic relations rather than anti-symmetric ones.
\iffalse
\be
C{(ij|kl)}-C{(ij|lk)} =0  \qquad \qquad N{(ij|kl)}-N{(ij|lk)} =0 
\ee
Since the symmetric color factors, $C^{\text{dd}}_g$, and kinematic numerators, $\tilde{N}^{\text{dd}}_g$, obey the same algebraic relations, gauge theories that are color-dual in the symmetric-sense can similarly be double copied by making the replacement $C^{\text{dd}}_g\rightarrow \tilde{N}^{\text{dd}}_g$:
\be
\mathcal{A} = \sum_{g}\frac{C^{\text{dd}}_gN^{\text{dd}}_g}{D_g} \quad \Rightarrow \quad \mathcal{M} = \sum_{g}\frac{\tilde{N}^{\text{dd}}_gN^{\text{dd}}_g}{D_g}
\ee
\fi
When a gauge theory can be expressed in a form such that the color and kinematic factors obey the same algebraic relations, we call such theories {color-dual}.   For ordered amplitudes  the ability to find kinematic numerators that obey antisymmetry and Jacobi relations is one-to-one with the ordered amplitudes satisfying KK and BCJ relations, respectively.

An important consequence of color-kinematics duality is the  ability to double-copy~\cite{BCJ} color-dual numerators with each other preserving simultaneously factorization of ordered-amplitudes as well as gauge-invariance.  Because cubic-graph color-weights are not linearly independent, gauge-invariance of the full amplitude is encoded in algebraic relation between the color factors. Thus, equipped with a set of color-dual numerators, $\tilde{N}_g$, that obey the same algebraic relations as $C_g$, we can make the simple replacement, $C_g \rightarrow \tilde{N}_g$, giving a new gauge invariant amplitude, $\mathcal{M}$, as follows:
\be\label{fullAmpCubicDouble}
\mathcal{M}_{n,L} = \int \prod_{i=1}^L\frac{d^D l_i}{(2\pi)^D} \sum_{g\in \Gamma^{(3)}_{n,L}} \frac{1}{S_g}\frac{\tilde{N}_gN_g}{D_g}
\ee
Note that $\mathcal{M}$ is colorless. It satisfies manifest gauge invariance on the $N_g$ side of the double copy due to the algebraic properties of $\tilde{N}_g$. In the event that $\tilde{N}_g$ also belongs to a vector theory, then $\mathcal{M}$ describes amplitudes in a gravitational theory. In such a case the gauge invariance of constituent $\mathcal{A}$ and $\tilde{\mathcal{A}}$ conspire to generate linearized diffeomorphism invariance.  As long as both kinematic factors obey the same algebraic constraints as the color algebra, this double copy construction works for integrands at the multi-loop level since it globally manifests color-kinematics on all possible unitarity cuts.

Throughout the text we will reserve $\mathcal{M}$ to denote double-copy amplitudes.  As shorthand, when the kinematic weights of two theories, $X$ and $Y$, participate in the double copy construction of $\mathcal{M}^{XY}$ we will use the outer-product to mean double copy construction as in,
\be
\mathcal{M}^{\text{XY}} = X \otimes Y \equiv \int \prod_{i=1}^L\frac{d^D l_i}{(2\pi)^D} \sum_{g\in \Gamma^{(3)}_{n,L}} \frac{1}{S_g}\frac{N^X_gN^Y_g}{D_g}\,.
\ee
At tree-level, such double-copy construction in the adjoint case can be understood equivalently \needRefs in terms of a KLT or momentum kernel matrix $S(a|b)$ and a BCJ spanning set of ordered amplitudes for each of th the $X$ and $Y$ theories,
\be
 X \otimes Y = \sum_{a,b \in S_{n-3}(2,\ldots,n-2)} A^X(1,\{a\},n,n-1) S(a|b) A^Y(1,\{b\},n-1,n)\,,
\ee
where the sum runs over both sets $a$, $b$ of $(n-3)!$ distinct ordered ampltidues.  No such momentum kernel has been constructed yet for the symmetric case.

@@

When working with boson kinematics in arbitrary dimensions, we will employ formal Lorentz covariant polarizations, $\varepsilon^\mu_a$, and momenta, $k^\mu_a$. The mostly minus signature will be used throughout. The only restrictions we will place on physical momenta and polarizations are the standard on-shell constraints: momentum conservation and the null-momenta condition,
\be
\sum_{a} k_a^\mu =0 \qquad \qquad k_a^2 =0
\ee
Furthermore, we will take external polarizations $\varepsilon_a$ formal and transverse throughout:
\be
\varepsilon_a \cdot k_a = 0
\ee
Dimensions will generically be away from four, $D=4-2\epsilon$, with $\epsilon$ arbitrary.

>>>>

\documentclass[11pt,letter]{article}
\usepackage{jheppub}
\usepackage{amsmath,amssymb}
\usepackage[dvipsnames]{xcolor}
%\usepackage{hyperref}
\usepackage{xspace}
\usepackage{ifdraft}
\usepackage{epstopdf}
\usepackage{slashed}
\usepackage{diagbox}
\usepackage{colortbl}
\usepackage{shuffle}
\usepackage{bbold}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{stackengine}
\def\RA{\rlap{\scalebox{1.6}{$\leftrightarrow$}}}
\def\DA{\smash{\bclap{\scalebox{1.6}{$\downarrow$}}}}
\def\mystrut{\rule[-2ex]{0ex}{6ex}}

\usepackage{tikz-feynman}

\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\usetikzlibrary{external}  
\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\tikzexternalize[                     %% Activate externalization
 prefix=images/,                    %% Avoid cluttering the directory
 system call={                       %% Use lualatex in system call
    pdflatex \tikzexternalcheckshellescape -halt-on-error -shell-escape -interaction=batchmode -jobname="\image" "\texsource" || rm "\image.pdf"
 },
]

\tikzset{graviton/.style={decorate, decoration={snake, amplitude=.4mm, segment length=1.5mm, pre length=.5mm, post length=.5mm}, double}}

\iffalse
\usepackage[framemethod=default]{mdframed}
\newmdenv[skipabove=7pt,
skipbelow=7pt,
rightline=false,
leftline=false,
topline=false,
bottomline=false,
backgroundcolor=gray!10,
linecolor=gray,
innerleftmargin=5pt,
innerrightmargin=5pt,
innertopmargin=5pt,
innerbottommargin=5pt,
leftmargin=0cm,
rightmargin=0cm,
linewidth=4pt]{eBox}
\fi

\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={funRed},
    citecolor={funBlue},
    urlcolor={funBlue}
}

@@

\definecolor{funRed}{RGB}{139,1,0}
\definecolor{funBlue}{RGB}{1,76,125}
\definecolor{cutred}{RGB}{219,56,49}
\definecolor{hgreen}{RGB}{25,176,146}
\definecolor{hgreen1}{RGB}{175,230,175}
\definecolor{hblue}{RGB}{52,152,219}
\definecolor{hbluedark}{RGB}{36, 106, 160}
\definecolor{hblue1}{RGB}{255,255,166}
\definecolor{hred}{RGB}{139,1,0}
\definecolor{hreddark}{RGB}{151, 58, 81}
\definecolor{hred1}{RGB}{255,155,155}
\definecolor{cutred}{RGB}{219,56,49}
\definecolor{hgrey4}{RGB}{75,75,75}
\definecolor{hgrey5}{RGB}{50,50,50}
\definecolor{hgrey3}{RGB}{100,100,100}
\definecolor{hgrey}{RGB}{125,125,125}
\definecolor{hgrey2}{RGB}{125,125,125}
\definecolor{hgrey1}{RGB}{150,150,150}
\definecolor{hgrey0}{RGB}{175,175,175}
\definecolor{hgreyLight}{RGB}{242,242,242}
\definecolor{darkgreen}{RGB}{59,126,108}

@@

\def\tree{{\rm tree}}
\def\pol{\varepsilon}
\def\Tr{\, {\rm Tr}}
\def\tr{\, {\rm tr}}
\def\eps{\varepsilon}
\def\e{\varepsilon}
\def\ep{\varepsilon}
\def\SYM{MSYM}
\def\nn{\nonumber}
\def\sec#1{section~\ref{#1}}
\def\eqn#1{eq.~(\ref{#1})}
\def\Eqn#1{Equation~(\ref{#1})}
\def\eqns#1#2{eqs.~(\ref{#1}) and~(\ref{#2})}
\def\Eqns#1#2{Eqs.~(\ref{#1}) and~(\ref{#2})}
\def\Figref#1{Fig.~\ref{#1}}
\def\secref#1{section~\ref{#1}}
\def\Neqfour{{{\cal N}=4}}
\def\NeqFour{{{\cal N}=4}}
\def\Neqeight{{{\cal N}=8}}
\def\NeqEight{{{\cal N}=8}}
\def\NeqOne{{{\cal N}=1}}
\def\Fact{{\cal F}}
\def\f{\widetilde f}

\def\Ksl{{\s K}}
\def\ksl{\s{k}}
\def\Perm{{\cal P}}
\def\M{{\cal M}}
\def\ve{\varepsilon}
\def\tlambda{{\tilde\lambda}}
\def\MHVbar{$\overline{\hbox{MHV}}$}
\def\NMHVbar{$\overline{\hbox{NMHV}}$}
\def\P{{\rm P}}
\def\NHP{{\rm NHP}}
\def\mud{\lambda}
\def\bowtie{{\rm bow\mbox{-}tie}}
\newcommand{\cred}{\bf \color{red}}
 \newcommand{\cblue}{\color{blue}}
 \definecolor{MattOrange}{rgb}{1.0,0.4,0.2}
\newcommand{\cob}{ \bf \color{MattOrange}}
\newcommand{\andd}{\ , \quad \text{and}  \quad}
\newcommand{\forr}{\ , \quad \text{for}  \quad}

@@

\title{Even-point Multi-loop Unitarity and its Applications: Exponentiation, Anomalies and Evanescence}

@@

\section{Introduction}\label{sec:intro}

Recent decades have seen  significant advances in our ability to compute scattering amplitudes to higher orders in perturbation theory. At the epicenter of this explosion in the generation of sharp $S$-matrix data is the unitarity method \cite{UnitarityMethod, Fusing, BDKUniarityReview}, which bypasses the standard Feynman diagram approach by constructing higher order amplitudes directly from lower order on-shell information. At one-loop, the unitarity method allows one to extract amplitudes by directly computing a series of unitarity cuts \cite{Forde:2007mi}. 
On-shell methods have furthermore unveiled novel amplitude-level structure, like the  duality between color and kinematics \cite{BCJ} and associated double-copy construction \cite{BCJLoop}, which are frequently obscured by the traditional classical Lagrangian description of the theories. 
While much of the research in perturbative calculations thus far has focused on gauge theory and gravity, recent literature has investigated to what extent the $S$-matrix of effective field theory is constrained by on-shell data \cite{Adams:2006sv,Cheung:2014dqa,Cheung:2015ota,Cheung:2016drk,Cheung:2018oki,Low:2019ynd,Carrasco:2019yyn,Arkani-Hamed:2020blm,Carrasco:2021ptp,Chi:2021mio,Bonnefoy:2021qgu,Carrasco:2022lbm,Carrasco:2022sck,Green:2022slj,Pavao:2022kog,Chen:2022shl,Chen:2023dcx,Brown:2023srz}.

@@

\section{Review} 
\label{sec:Review}
\subsection{Color-dressed and ordered amplitudes}
Here we provide an overview of the amplitudes nomenclature and organizational principles we use throughout the work. When working with scattering amplitudes, $\mathcal{A}$, in an on-shell framework it is convenient to introduce the following graphical description at general multiplicity, $n$, and loop order, $L$,
\begin{equation}
\label{fullAmp}
\mathcal{A}_{n,L} = \int \prod_{i=1}^L\frac{d^D l_i}{(2\pi)^D} \sum_{g} \frac{1}{S_g}\frac{\mathcal{N}_g}{D_g}
\end{equation}
As written above, $S_g$ are the internal symmetry factors for a Feynman diagram, $g$, the propagator structure is captured by the denominator function, $D_g$, and the theory-dependent interaction vertices determine numerator functions, $\mathcal{N}_g$. As written the numerator functions carry all non-propagator kinematic information and any color-data that would be encoded in the interaction Feynman rules of the theory. By color-data we mean the typical weighting of feynman diagrams by the representation of the particles described.  Most of the theories we describe are defined in the adjoint, so the color-data involves the usual dressing of vertices with  antisymmetric structure constants, $f^{abc}$.  Although, as we will see, it will be useful to describe certain counterterms in terms of symmetric $d^{abc}$ color-weights,

Fo  gauge theories, it is often useful to further decompose the full, or color-dressed, amplitudes of \eqn{fullAmp} into purely kinematic building blocks where the color information is stripped away. For example, at tree level, gauge theory $n$-point amplitudes can be re-expressed in terms of a trace basis decomposition as follows:
\begin{equation}
\mathcal{A}_{n,\text{tree}} = \sum_{\sigma \in S^{n-1}} \text{Tr}(T^1 T^{\sigma(2)}\cdots T^{\sigma(n)}) A_n(1,\sigma(2),...,\sigma(n))
\end{equation}
where the sum is taken over the inequivalent orderings of group theory generator traces. A similar decomposition applies to one-loop gauge theory amplitudes, which we will describe in \sect{sec:Loops}. The color-ordered functions, $A_n$, are called {partial amplitudes}, as they only contain on-shell information for a particular color ordering. However, like full amplitudes, they factor on poles to lower-point partial amplitudes. For vector theories partial amplitudes are independently gauge invariant since they weight linearly independent color traces.

When expressed in this form, many gauge theories reveal hidden redundancy that is obscured by the color-dressed amplitude formulation of \eqn{fullAmp}. The $(n-1)!$ partial amplitudes for a large catalog \cite{BCJreview} of adjoint gauge theories, including Yang-Mills and NLSM, are linearly related to a smaller set of basis amplitudes. The first such set of partial amplitude relations identified by Kleiss and Kluijf \cite{Kleiss:1988ne}, known formally as KK relations, relates amplitudes different color orderings and signature to a basis of $(n-2)!$ partial amplitudes:
\begin{equation}
\label{KK}
A(1,\alpha,n,\beta)=(-1)^{|\alpha|}\sum_{\sigma\in \alpha \shuffle \beta^T} A(1,\sigma,n)
\end{equation}
where $\beta^T$ denotes the inverse ordering of $\beta$, and $\shuffle$ is the shuffle product that runs over ordered permutations of $\alpha$ and $\beta^T$. More recently, Bern, Johansson and one of the authors, (BCJ) further identified a set of kinematic relations that reduced the size of $n$-point partial amplitude basis down to $(n-3)!$ via BCJ relations \cite{BCJ}:
\begin{equation}\label{BCJ}
\sum_{i=2}^{n-1}k_1\cdot(k_2+\cdots +k_i) A(2,...,i,1,...,n)=0
\end{equation}
Underlying both sets of amplitude relations are hidden graphical  principles that manifest the amplitude level redundancy of \eqn{KK} and \eqn{BCJ}, which we  describe in the next section. Before moving on, we emphasize that everything described thus far applies in arbitrary spacetime dimensions. As such, the aforementioned nomenclature and organization are well suited for constructing loop integrands compatible with  dimensional regularization.
\label{subsec:AmpReview}

\subsection{Color-Kinematics Duality and the Double-Copy}
\label{subsec:DCReview}
To apprehend this graph-based structure for color-dressed theories, we first rewrite the full amplitude of \eqn{fullAmp} as a sum over {cubic graphs}, as follows:
\begin{equation}\label{fullAmpCubic}
\mathcal{A}_{n,L} = \int \prod_{i=1}^L\frac{d^D l_i}{(2\pi)^D} \sum_{g\in \Gamma^{(3)}_{n,L}} \frac{1}{S_g}\frac{C_gN_g}{D_g}
\end{equation}
where $N_g$ and $C_g$ are the kinematic numerators and color factors, respectively, and $ \Gamma^{(3)}_{n,L}$ denotes the set of cubic diagrams at $n$-point $L$-loop. In this formulation of the amplitude, contact diagrams are absorbed into cubic graphs by multiplying by inverse propagators. The procedure of assigning contact diagrams to cubic graphs is referred to as {generalized gauge freedom} \cite{BCJ}. In practice, there are many different ways to assign contact diagrams to cubic graphs.

The gauge theories that we study in this paper have sufficient generalized gauge freedom to assign contact diagrams to kinematic numerators such that at every multiplicity at tree level one can write the amplitude in terms of $N_g$ that obey the same algebraic relations as the color factors, $C_g$. For color factors that are composed of adjoint structure constants, $f^{abc}$ they must be anti-symmetric around vertices, and be related by Jacobi relations about each edge.   For $SU(N_c)$ we can take $f^{abc} \propto \text{Tr}(T^a[T^b,T^c])$ for arbitrary representation $T$ of the gauge group.
\iffalse
Explicitly, given the internal edge of a cubic graph, $\mathcal{G}{(ij|kl)}$, described functionally in terms of external edges,
\begin{equation}
\mathcal{G}{(ij|kl)}=\threeTree{$j$}{$k$}{$l$}
\end{equation}
we can construct color factors, $C(ij|kl)$, and kinematic factors, $N(ij|kl)$, that obey the following set of functional relations on all vertices of internal edges:
\begin{itemize}
\item \textbf{Antisymmetry Relations} $\Leftrightarrow$ \textbf{KK Relations}
\begin{equation}
C{(ij|kl)}+C{(ij|lk)} =0  \qquad \qquad N{(ij|kl)}+N{(ij|lk)} =0 
\end{equation}
\item \textbf{Jacobi Relations} $\Leftrightarrow$ \textbf{BCJ Relations}
\begin{equation}
C{(ij|kl)}+C{(il|jk)} +C{(ik|lj)} =0  \qquad N{(ij|kl)}+N{(il|jk)} +N{(ik|lj)} =0 
\end{equation}
\end{itemize}
\fi
We will also consider generalization beyond the adjoint to color-weights that can include $d^{abc}$ and associated algebraic relations.   For $SU(N_c)$ we can take $d^{abc} \propto Tr(T^a\{T^b,T^c\})$.  Symmetric-structure color obey symmetric algebraic relations rather than anti-symmetric ones.
\iffalse
\begin{equation}
C{(ij|kl)}-C{(ij|lk)} =0  \qquad \qquad N{(ij|kl)}-N{(ij|lk)} =0 
\end{equation}
Since the symmetric color factors, $C^{\text{dd}}_g$, and kinematic numerators, $\tilde{N}^{\text{dd}}_g$, obey the same algebraic relations, gauge theories that are color-dual in the symmetric-sense can similarly be double copied by making the replacement $C^{\text{dd}}_g\rightarrow \tilde{N}^{\text{dd}}_g$:
\begin{equation}
\mathcal{A} = \sum_{g}\frac{C^{\text{dd}}_gN^{\text{dd}}_g}{D_g} \quad \Rightarrow \quad \mathcal{M} = \sum_{g}\frac{\tilde{N}^{\text{dd}}_gN^{\text{dd}}_g}{D_g}
\end{equation}
\fi
When a gauge theory can be expressed in a form such that the color and kinematic factors obey the same algebraic relations, we call such theories {color-dual}.   For ordered amplitudes  the ability to find kinematic numerators that obey antisymmetry and Jacobi relations is one-to-one with the ordered amplitudes satisfying KK and BCJ relations, respectively.

An important consequence of color-kinematics duality is the  ability to double-copy~\cite{BCJ} color-dual numerators with each other preserving simultaneously factorization of ordered-amplitudes as well as gauge-invariance.  Because cubic-graph color-weights are not linearly independent, gauge-invariance of the full amplitude is encoded in algebraic relation between the color factors. Thus, equipped with a set of color-dual numerators, $\tilde{N}_g$, that obey the same algebraic relations as $C_g$, we can make the simple replacement, $C_g \rightarrow \tilde{N}_g$, giving a new gauge invariant amplitude, $\mathcal{M}$, as follows:
\begin{equation}\label{fullAmpCubicDouble}
\mathcal{M}_{n,L} = \int \prod_{i=1}^L\frac{d^D l_i}{(2\pi)^D} \sum_{g\in \Gamma^{(3)}_{n,L}} \frac{1}{S_g}\frac{\tilde{N}_gN_g}{D_g}
\end{equation}
Note that $\mathcal{M}$ is colorless. It satisfies manifest gauge invariance on the $N_g$ side of the double copy due to the algebraic properties of $\tilde{N}_g$. In the event that $\tilde{N}_g$ also belongs to a vector theory, then $\mathcal{M}$ describes amplitudes in a gravitational theory. In such a case the gauge invariance of constituent $\mathcal{A}$ and $\tilde{\mathcal{A}}$ conspire to generate linearized diffeomorphism invariance.  As long as both kinematic factors obey the same algebraic constraints as the color algebra, this double copy construction works for integrands at the multi-loop level since it globally manifests color-kinematics on all possible unitarity cuts.

Throughout the text we will reserve $\mathcal{M}$ to denote double-copy amplitudes.  As shorthand, when the kinematic weights of two theories, $X$ and $Y$, participate in the double copy construction of $\mathcal{M}^{XY}$ we will use the outer-product to mean double copy construction as in,
\begin{equation}
\mathcal{M}^{\text{XY}} = X \otimes Y \equiv \int \prod_{i=1}^L\frac{d^D l_i}{(2\pi)^D} \sum_{g\in \Gamma^{(3)}_{n,L}} \frac{1}{S_g}\frac{N^X_gN^Y_g}{D_g}\,.
\end{equation}
At tree-level, such double-copy construction in the adjoint case can be understood equivalently \needRefs in terms of a KLT or momentum kernel matrix $S(a|b)$ and a BCJ spanning set of ordered amplitudes for each of th the $X$ and $Y$ theories,
\begin{equation}
 X \otimes Y = \sum_{a,b \in S_{n-3}(2,\ldots,n-2)} A^X(1,\{a\},n,n-1) S(a|b) A^Y(1,\{b\},n-1,n)\,,
\end{equation}
where the sum runs over both sets $a$, $b$ of $(n-3)!$ distinct ordered ampltidues.  No such momentum kernel has been constructed yet for the symmetric case.

@@

When working with boson kinematics in arbitrary dimensions, we will employ formal Lorentz covariant polarizations, $\varepsilon^\mu_a$, and momenta, $k^\mu_a$. The mostly minus signature will be used throughout. The only restrictions we will place on physical momenta and polarizations are the standard on-shell constraints: momentum conservation and the null-momenta condition,
\begin{equation}
\sum_{a} k_a^\mu =0 \qquad \qquad k_a^2 =0
\end{equation}
Furthermore, we will take external polarizations $\varepsilon_a$ formal and transverse throughout:
\begin{equation}
\varepsilon_a \cdot k_a = 0
\end{equation}
Dimensions will generically be away from four, $D=4-2\epsilon$, with $\epsilon$ arbitrary.
